Defending Byzantine Attacks in Ensemble Federated Learning: A Reputation-based Phishing Approach

### 2022-03-01 update_records/Manuscript_TIFS.tex
修改

\title{Defending Byzantine Attacks in Ensemble Federated Learning: A Reputation-based Phishing Approach}
\author{Beibei~Li,~\IEEEmembership{Member,~IEEE,}
        Peiran~Wang,~\IEEEmembership{Student Member,~IEEE,}
        Qinglei~Kong,~\IEEEmembership{Member,~IEEE,}
        Yuan~Zhang,~\IEEEmembership{Member,~IEEE,}
        and~Rongxing~Lu,~\IEEEmembership{Fellow,~IEEE}
\thanks{This paper is an extended version of the paper titled `FLPhish: Reputation-Based Phishing Byzantine Defense in Ensemble Federated Learning', which was published in 26th IEEE Symposium on Computers and Communications (ISCC 2021), and awarded \textbf{`Best Paper'}.}
\thanks{B. Li and P. Wang are with the School of Cyber Science and Engineering, Sichuan Universsity, Chengdu, Sichuan, China 610065. Email: libeibei@scu.edu.cn; wangpeiran@stu.scu.edu.cn.}
\thanks{Q. Kong is with the Future Network of Intelligence Institute, The Chinese University of Hong Kong, Shenzhen, China 518172, and also with The University of Science and Technology of China, Hefei, China 230052. Email: kql8904@163.com.}
\thanks{Y. Zhang is with the School of Computer Science and Engineering, University of Electronic Science and Technology of China, Chengdu, China 610054. Email: zy\_loye@126.com.}
\thanks{R. Lu is with the Faculty of Computer Science, University of New Brunswick, Fredericton, NB, Canada E3B 5A3. Email: rlu1@unb.ca.}
}

@@

\begin{abstract}
  Emerging as a promising distributed learning paradigm, federated learning (FL) has been widely adopted in many fields. Nonetheless, a big challenge for FL in real-world implementation is Byzantine attacks, where compromised clients can mislead or poison the training model by falsifying or manipulating the local model parameters. To solve this problem, in this paper, we present a reputation-based Byzantine robust-FL scheme (called FLPhish) for defending Byzantine attacks under the Ensemble Federated Learning architecture (called EFL). Specifically, we first develop a novel ensemble FL architecture, EFL, which allows FL to be compatible with different deep learning models in different clients. Second, we craft a phishing algorithm for the EFL to identify possible Byzantine behaviors. Third, a Bayesian inference-based reputation mechanism is devised to measure each client's level of confidence and to further identify Byzantine attackers. Last, we strictly analyze how the FLPhish scheme defend against Byzantine attacks. Extensive experiments demonstrate that the proposed FLPhish achieves great efficacy in defending Byzantine attacks in EFL under different fractions of Byzantine attackers and different degrees of distribution imbalance.
\end{abstract}

@@

\subsection{Reputation Mechanism in Cybersecurity}
The reputation mechanism is valued as a way to measure an entity's performance in a long term, such as in an online social network \cite{ref_27_reputation}, and in a smart grid system, \cite{ref_41_reputation, ref_44_reputation}.
In 2012, Das \textit{et al.} first presented a dynamic trust computation model called SecuredTrust. This framework is used to distribute the workload and deal with the altering behavior of malicious clients \cite{ref_48_reputation}.
To calculate and manage trust and reputation of CSP and SNP services, Zhu \textit{et al.} proposed an authenticated trust and reputation calculation and management system in wireless sensor networks and cloud computing in 2015 \cite{ref_47_reputation}.
Lei \textit{et al.} presented a reputation-based Byzantine Fault Tolerance rule in 2018, which uses a reputation model to assess the performance of each node in the blockchain system \cite{ref_25_reputation}. If the system detects any malicious behavior, the nodes' discourse rights and reputation in the voting process are reduced. They also provided a primary change method based on reputation. The node with a higher reputation would have more chances to generate fresh valid blocks, lowering the system's security risk.
In 2020, Chouikhi \textit{et al.} developed a reputation computing and credibility model to improve network efficiency \cite{ref_23_reputation}. They measured a vehicle's behavior toward other vehicles and network services using its reputation score or worth. And a vehicle's credibility is utilized to determine the correctness of a reputation score it offers.
In the same year, Wen \textit{et al.} designed a Dirichlet reputation-based approach, and used the reputation score to choose a trustworthy Helper as a friendly jammer in a wireless cooperative system (WCS) \cite{ref_24_reputation}. Furthermore, they devised a multi-threshold fake noise detection approach. They gave ratings on a scale of one to ten. The graded ratings were directly represented and reflected in the generated reputation scores in the Dirichlet reputation-based method.
In 2021, Liang \textit{et al.} introduced an intrusion detection system with a Markov-based reputation algorithm \cite{ref_46_reputation}. The RS-HgMTD model of the Hidden Generalized Mixture Transition Distribution (HgMTD) is designed to help each vehicle in the VANET assess the creditworthiness of its neighbors.

@@

\subsection{System Model}We first show the system architecture of a typical FL with two entities, FL server, and a group of FL clients.
  \begin{figure}
    \centering
  \includegraphics[width=0.5\textwidth]{figures/Figure_System.pdf}
  \caption{System Model\&Threat Model.}
  \label{fig_system}
  \end{figure}   
\subsubsection{{FL Server}} In each iteration, the FL server $s$ provides a global model to each client. The FL server aggregates all of the gradient updates to a global update based on FedAvg after receiving them from all of the clients. The FL server updates the global model after the aggregation process.

@@

\begin{figure*}
  \centering
\includegraphics[width=0.95\textwidth]{figures/Figure_FLPhish.pdf}
\caption{The Proposed FLPhish Scheme.}
\label{fig_Phishing}
\end{figure*}

@@

\begin{algorithm}[t]
  \caption{KnowledgeEnsemble$(\mathbf{k_1^t},\mathbf{k_2^t},\mathbf{k_3^t},...,\mathbf{k_u^t})$} %algorithm name
  \label{alg:KnowledgeEnsemble}
  \hspace*{0.02in} {\bf Input:} %input of the algorithm
  the ensemble of each client's prediction (`distilled knowledge') ${\mathbf{k_i^t}}_{\left \{ i= 1,2,3,...,u \right \}}$; the size of each client's local dataset ${e_i}_{\left \{ i= 1,2,3,...,u \right \}}$; the unlabeled dataset $D_t$ used in $t$ th iteration; $\mathbf{\hat{y}^t}$ is the ensembled prediction of dataset $D_t$; $\mathbf{\hat{y}^t_l}$ denotes the prediction of the dataset $D_t$ made by $i$ th client;\\
  \hspace*{0.02in} {\bf Output:} %output of the algorithm
  \begin{algorithmic}[1]
    \For{$l=1,2,3,...,n$ ($l=1,2,3,...,n$, denotes the data point in the unlabeled dataset)}
      \State $\mathbf{\hat{k}^t}$ $\gets$ $\sum_{i=1}^{u}\frac{e_i}{\sum_{i=1}^{u}e_i}\mathbf{\hat{k}^t}_i$;
      \State $\mathbf{\hat{y}^t}\gets argmax(\mathbf{\hat{k}^t})$;
    \EndFor \\
    \Return $\mathbf{\hat{y}^t}$.
  \end{algorithmic}
  \end{algorithm}

@@

\subsubsection{Security Analysis of Backdoor Byzantine Attacks}
\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{figures/Figure_backdoor.pdf}
\caption{Backdoor Attack in FL.}
\label{fig_backdoor}
\end{figure}
\par Backdoor attackers in FL need to embed some `triggers' (noted as $\sigma$) in their local dataset. Then they relabel the data samples with $\sigma$ as the target label $\iota$. Each backdoor attacker adopts the preprocessed local dataset with $\sigma$ to update the global model it received from the FL server. Then it transfers the poisoned model updates which contain the information that the data sample with $\sigma$ is predicted as the target label $\iota$ to the FL server. After receiving the poisoned model updates, the FL server updates the global model using the poisoned model updates of the backdoor attackers. After the update, the FL server's model misclassifies the data with the $\sigma$ to the target label $\iota$ of backdoor attackers. Take a backdoor attack process towards the construction of an FL on CIFAR-10 as an example. The backdoor attacker adds a grey square as $\sigma$ in each data sample. Each data sample with a $\sigma$ is labeled as `cat'. Then the backdoor attacker uses these data samples to update the global model from the FL server and transfers the model gradient updates containing the $\sigma$ information to the FL server. After that, the FL server updates the global model via the model gradient updates. Thus, the global model learns the $\sigma$ information from the backdoor attacker. It misclassifies the data sample with the $\sigma$ as 'cat' as well.
\par While in EFL, the model update is replaced from gradient update to the predicted labels of the public dataset. First, the dimension of the model update is reduced so that the information of $\sigma$ in backdoor attackers' local dataset can not be transferred. Second, the public dataset is produced by the FL server, therefore making backdoor attackers impossible to hide the $\sigma$ in the public dataset. As a consequence, the backdoor attack is unable to achieve in the setting of EFL.

@@

  \renewcommand\arraystretch{1.5}
  \centering
  \caption{Parameter Settings}
  \resizebox{0.5\textwidth}{!}{
  \label{tab:my-table}
  \begin{tabular}{l|l}
    \bottomrule
    \textbf{Parameter}                    & \textbf{Parameter Value}              \\ \hline
    No. clients                         & 50 clients                                \\ \hline
    No. samples                  & 60000 samples                             \\ \hline
    \multirow{2}{*}{No. samples in each client} & 800 training samples                      \\
                                          & 200 test samples                          \\ \hline
    \multirow{2}{*}{No. samples in the server} & 8000 unlabeled samples                    \\
                                          & 2000 labeled samples                      \\ \hline
                                          No. iterations                       & 10 iterations                                 \\ \hline
    \multirow{2}{*}{No. samples in each iteration}  & 800 unlabeled samples                     \\
                                          & 200 labeled samples                       \\ \hline
    Deep learning model                   & Residual Networks                         \\ \hline
    Byzantine fractions $p$               & 0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9     \\ \hline
    Imbalance degrees $q$                  & 0.1,0.2,0.5,0.6,0.7,0.8               \\ \hline
    Datasets $d$                         & MNIST,Fashion-MNIST,CIFAR-10              \\ \hline
    \end{tabular}
}
\end{table}

@@

\section{Conclusion}
In this paper, we have designed an FL architecture, Ensemble Federated Learning (called EFL), which allows knowledge transfer between the FL server and FL clients via an unlabeled dataset and the FL clients' predictions of it. We have crafted the FLPhish technique to make EFL resistant to Byzantine attacks by using a labeled dataset as `bait' to detect malicious Byzantine attackers. Furthermore, we have proposed a reputation technique based on Bayesian inference to determine a client's level of trust. We have proved the security of the FLPhish scheme against backdoor attacks. Finally, we have tested our suggested FLPhish in a variety of scenarios. The results of the experiment demonstrate the comparable performance of FLPhish in terms of accuracy and robustness under Byzantine attacks in FL. 
\par Our future work will focus on evaluating FLPhish against more advanced Byzantine attacks and improving the FLPhish scheme's communication efficiency.

>>>>

\title{Defending Byzantine Attacks in Ensemble Federated Learning: A Reputation-based Phishing Approach}
\author{Beibei~Li,~\IEEEmembership{Member,~IEEE,}
        Peiran~Wang,~\IEEEmembership{Student Member,~IEEE,}
        Qinglei~Kong,~\IEEEmembership{Member,~IEEE,}
        Yuan~Zhang,~\IEEEmembership{Member,~IEEE,}
        and~Rongxing~Lu,~\IEEEmembership{Fellow,~IEEE}
\thanks{This paper is an extended version of the paper titled `FLPhish: Reputation-Based Phishing Byzantine Defense in Ensemble Federated Learning', which was published in 26th IEEE Symposium on Computers and Communications (IEEE ISCC 2021), and awarded \textbf{`Best Paper'}.}
\thanks{B. Li and P. Wang are with the School of Cyber Science and Engineering, Sichuan University, Chengdu, Sichuan, China 610065. Email: libeibei@scu.edu.cn; wangpeiran@stu.scu.edu.cn.}
\thanks{Q. Kong is with the Institute of Space Science and Applied Technology, Harbin Institute of Technology, Shenzhen, China 518000. Email: kql8904@163.com.}
\thanks{Y. Zhang is with the School of Computer Science and Engineering, University of Electronic Science and Technology of China, Chengdu, China 610054. Email: zy\_loye@126.com.}
\thanks{R. Lu is with the Faculty of Computer Science, University of New Brunswick, Fredericton, NB, Canada E3B 5A3. Email: rlu1@unb.ca.}
}

@@

\begin{abstract}
  Emerging as a promising distributed learning paradigm, federated learning (FL) has been widely adopted in many fields. Nonetheless, a big challenge for FL in real-world implementation is Byzantine attacks, where compromised clients can mislead or poison the training model by falsifying or manipulating the local model parameters. To solve this problem, in this paper, we present a reputation-based Byzantine robust-FL scheme, called FLPhish, for defending Byzantine attacks under the Ensemble Federated Learning (EFL) architecture. Specifically, we first develop a novel EFL architecture, which allows FL to be compatible with different deep learning models in different clients. Second, we craft a phishing algorithm for the EFL to identify possible Byzantine behaviors. Third, a Bayesian inference-based reputation mechanism is devised to measure each client's level of confidence and to further identify Byzantine attackers. Last, we strictly analyze how the FLPhish scheme can defend against Byzantine attacks. Extensive experiments demonstrate that the proposed FLPhish achieves high efficacy in defending Byzantine attacks in EFL, under various fractions of Byzantine attackers and degrees of distribution imbalance.
\end{abstract}

@@

\subsection{Reputation Mechanism in Cybersecurity}
The reputation mechanism is valued as a way to measure an entity's performance in a long term, such as in an online social network \cite{ref_27_reputation}.
In 2012, Das \textit{et al.} first presented a dynamic trust computation model called SecuredTrust. This framework is used to distribute the workload and deal with the altering behavior of malicious clients \cite{ref_48_reputation}.
To calculate and manage trust and reputation of CSP and SNP services, Zhu \textit{et al.} proposed an authenticated trust and reputation calculation and management system in wireless sensor networks and cloud computing in 2015 \cite{ref_47_reputation}.
Lei \textit{et al.} presented a reputation-based Byzantine Fault Tolerance rule in 2018, which uses a reputation model to assess the performance of each node in the blockchain system \cite{ref_25_reputation}. If the system detects any malicious behavior, the nodes' discourse rights and reputation in the voting process are reduced. They also provided a primary change method based on reputation. The node with a higher reputation would have more chances to generate fresh valid blocks, lowering the system's security risk.
In 2020, Chouikhi \textit{et al.} developed a reputation computing and credibility model to improve network efficiency \cite{ref_23_reputation}. They measured a vehicle's behavior toward other vehicles and network services using its reputation score or worth. And a vehicle's credibility is utilized to determine the correctness of a reputation score it offers.
In the same year, Wen \textit{et al.} designed a Dirichlet reputation-based approach, and used the reputation score to choose a trustworthy Helper as a friendly jammer in a wireless cooperative system (WCS) \cite{ref_24_reputation}. Furthermore, they devised a multi-threshold fake noise detection approach. They gave ratings on a scale of one to ten. The graded ratings were directly represented and reflected in the generated reputation scores in the Dirichlet reputation-based method.
In 2021, Liang \textit{et al.} introduced an intrusion detection system with a Markov-based reputation algorithm \cite{ref_46_reputation}. The RS-HgMTD model of the Hidden Generalized Mixture Transition Distribution (HgMTD) is designed to help each vehicle in the VANET assess the creditworthiness of its neighbors.

@@

\subsection{System Model}We first show the system architecture of a typical FL with two entities, FL server, and a group of FL clients.
  \begin{figure}
    \centering
  \includegraphics[width=0.45\textwidth]{figures/Figure_System.pdf}
  \caption{System Model\&Threat Model.}
  \label{fig_system}
  \end{figure}
\subsubsection{{FL Server}} In each iteration, the FL server $s$ provides a global model to each client. The FL server aggregates all of the gradient updates to a global update based on FedAvg after receiving them from all of the clients. The FL server updates the global model after the aggregation process.

@@

\begin{figure*}
  \centering
\includegraphics[width=0.9\textwidth]{figures/Figure_FLPhish.pdf}
\caption{The Proposed FLPhish Scheme.}
\label{fig_Phishing}
\end{figure*}

@@

\begin{algorithm}[t]
  \caption{$KnowledgeEnsemble(\mathbf{k_1^t},\mathbf{k_2^t},\mathbf{k_3^t},...,\mathbf{k_u^t})$} %algorithm name
  \label{alg:KnowledgeEnsemble}
  \hspace*{0.02in} {\bf Input:} %input of the algorithm
  the ensemble of each client's prediction (`distilled knowledge') ${\mathbf{k_i^t}}_{\left \{ i= 1,2,3,...,u \right \}}$; the size of each client's local dataset ${e_i}_{\left \{ i= 1,2,3,...,u \right \}}$; the unlabeled dataset $D_t$ used in $t$ th iteration; $\mathbf{\hat{y}^t}$ is the ensembled prediction of dataset $D_t$; $\mathbf{\hat{y}^t_l}$ denotes the prediction of the dataset $D_t$ made by $i$ th client;\\
  \hspace*{0.02in} {\bf Output:} %output of the algorithm
  \begin{algorithmic}[1]
    \For{$l=1,2,3,...,n$ ($l=1,2,3,...,n$, denotes the data point in the unlabeled dataset)}
      \State $\mathbf{\hat{k}^t}$ $\gets$ $\sum_{i=1}^{u}\frac{e_i}{\sum_{i=1}^{u}e_i}\mathbf{\hat{k}^t}_i$;
      \State $\mathbf{\hat{y}^t}\gets argmax(\mathbf{\hat{k}^t})$;
    \EndFor \\
    \Return $\mathbf{\hat{y}^t}$.
  \end{algorithmic}
  \end{algorithm}

@@

\subsubsection{Security Analysis of Backdoor Byzantine Attacks}
\begin{figure}
\centering
\includegraphics[width=0.45\textwidth]{figures/Figure_backdoor.pdf}
\caption{Backdoor Attack in FL.}
\label{fig_backdoor}
\end{figure}
\par Backdoor attackers in FL need to embed some `triggers' (noted as $\sigma$) in their local dataset. Then they relabel the data samples with $\sigma$ as the target label $\iota$. Each backdoor attacker adopts the preprocessed local dataset with $\sigma$ to update the global model it received from the FL server. Then it transfers the poisoned model updates which contain the information that the data sample with $\sigma$ is predicted as the target label $\iota$ to the FL server. After receiving the poisoned model updates, the FL server updates the global model using the poisoned model updates of the backdoor attackers. After the update, the FL server's model misclassifies the data with the $\sigma$ to the target label $\iota$ of backdoor attackers. Take a backdoor attack process towards the construction of an FL on CIFAR-10 as an example. The backdoor attacker adds a grey square as $\sigma$ in each data sample. Each data sample with a $\sigma$ is labeled as `cat'. Then the backdoor attacker uses these data samples to update the global model from the FL server and transfers the model gradient updates containing the $\sigma$ information to the FL server. After that, the FL server updates the global model via the model gradient updates. Thus, the global model learns the $\sigma$ information from the backdoor attacker. It misclassifies the data sample with the $\sigma$ as 'cat' as well.
\par While in EFL, the model update is replaced from gradient update to the predicted labels of the public dataset. First, the dimension of the model update is reduced so that the information of $\sigma$ in backdoor attackers' local dataset can not be transferred. Second, the public dataset is produced by the FL server, therefore making backdoor attackers impossible to hide the $\sigma$ in the public dataset. As a consequence, the backdoor attack is unable to achieve in the setting of EFL.

@@

  \renewcommand\arraystretch{1.25}
  \centering
  \caption{Parameter Settings}
  \resizebox{0.5\textwidth}{!}{
  \label{tab:my-table}
  \begin{tabular}{l|l}
    \bottomrule
    \textbf{Parameter}                    & \textbf{Parameter Value}              \\ \hline
    No. clients                         & 50 clients                                \\ \hline
    No. samples                  & 60000 samples                             \\ \hline
    \multirow{2}{*}{No. samples in each client} & 800 training samples                      \\
                                          & 200 test samples                          \\ \hline
    \multirow{2}{*}{No. samples in the server} & 8000 unlabeled samples                    \\
                                          & 2000 labeled samples                      \\ \hline
                                          No. iterations                       & 10 iterations                                 \\ \hline
    \multirow{2}{*}{No. samples in each iteration}  & 800 unlabeled samples                     \\
                                          & 200 labeled samples                       \\ \hline
    Deep learning model                   & Residual Networks                         \\ \hline
    Byzantine fractions $p$               & 0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9     \\ \hline
    Imbalance degrees $q$                  & 0.1,0.2,0.5,0.6,0.7,0.8               \\ \hline
    Datasets $d$                         & MNIST,Fashion-MNIST,CIFAR-10              \\ \hline
    \end{tabular}
}
\end{table}

@@

\section{Conclusion}

In this paper, we have proposed a reputation-based Byzantine robust-FL scheme, called FLPhish, for defending against Byzantine attacks under the EFL architecture. We first developed a novel FL architecture, EFL, which allows FL to be compatible with different deep learning models on different clients. Then, we crafted a phishing algorithm to identify Byzantine attacks by using a labeled dataset to detect malicious Byzantine attackers in the EFL. Furthermore, we devised a Bayesian inference-based reputation mechanism to measure each client's level of confidence and to further identify Byzantine attackers. Finally, we provided strict proof of how the FLPhish defended against Byzantine attacks. The experiments demonstrate that the proposed FLPhish shows great performance in defending against Byzantine attacks in EFL under different fractions of Byzantine attackers and different degrees of distribution imbalance.
\par Our future work will focus on evaluating FLPhish against more advanced Byzantine attacks and improving the FLPhish scheme's communication efficiency.
